---
title: How to get data from Google trends API in R
author:
  - name:
      given: David
      family: Schoch
    email: david.schoch@gesis.org
    orcid: 0000-0003-2952-4812
    affiliations:
      - name: "GESIS - Leibniz Institute for the Social Sciences"
csl: apa.csl
format:
  html: default
  ipynb: default
license: CC BY
---


## Learning Objectives

By the end of this tutorial, you will be able to obtain data from [Google trends](https://trends.google.com)[^1] using the R package 
[gtrendsR](https://github.com/PMassicotte/gtrendsR)

[^1]:  Google trends shows how frequently a given search term is searched for, relative to the total search volume over a given time period.

## Target audience

This tutorial is targeted for people who want to retrieve data from Google trends

## Setting up the computational environment

The following R package is required:
```{r}
#| label: install
#| eval: false

install.packages("gtrendsR")
```
(*This tutorial is written with version `r packageVersion("gtrendsR")`.*)  

No further configuration is needed, but make sure to not send too many requests in a short period of time
because Google can decide to block you.

The package can be loaded with
```{r}
#| label: load

library(gtrendsR)
```

For data wrangling and visualization tasks we use the [tidyverse](https://www.tidyverse.org/).

```{r}
#| label: tidyverse
#| message: false
#| warning: false

# install.packages("tidyverse")
library(tidyverse)
```

## Social Science Usecase

Google Trends data can be a valuable resource for examining public interest and behavior patterns over time. For instance to
analyze public health awareness and behavior in response to disease outbreaks. The objective here is to understand how awareness and public interest in specific health issues, like COVID-19, influenza, or mental health topics, change over time and how these patterns correlate with real-world events or policy implementations.

## Getting data

The Package only has one function called `gtrends()` which does all the querying of Google Trends automatically. You can get help about its parameters with
```{r}
#| label: help
#| eval: false

?gtrends
```

Supplying just the keyword queries for global trend data of the last five years on a weekly basis.
```{r}
#| label: example1

covid_tbl <- gtrends(keyword = "covid")
```

The `trend_tbl` is a list of data frames with different granularity of the trend result.
```{r}
#| label: trend_object
names(covid_tbl)
```

`interest_over_time` includes the search relevance over time.
```{r}
#| label: plot_interest_over_time_covid
covid_tbl$interest_over_time |> 
    mutate(hits = as.numeric(gsub("<", "", hits))) |> 
    ggplot(aes(x = date, y = hits))+
    geom_line()

```

Several keywords can be specified to compare search trends.

```{r}
#| label: example2
#| cache: true

covax_tbl <- gtrends(keyword = c("covid","vaccine"))
```

```{r}
#| label: plot_interest_over_time_covax
covax_tbl$interest_over_time |> 
    mutate(hits = as.numeric(gsub("<","",hits))) |> 
    ggplot(aes(x = date, y = hits, col = keyword))+
    geom_line()

```

The data frame `related_queries` lists search queries that are most related to the provided keyword(s).

```{r}
#| label: plot_related_queries_covid

covid_tbl$related_queries |> 
    filter(related_queries == "top") |> 
    mutate(subject = as.numeric(subject)) |> 
    top_n(10,subject) |> 
    ggplot(aes(x = subject,y = reorder(value, subject))) + 
    geom_col()

```

## Other things to try

The examples above use default values for the parameters `geo` and `time` which 
return world-wide trends of the last 5 years. To restrict results to (a) specific location(s) use
2 letter country codes provided in `data("countries")` for the `geo` parameter. 

To adjust the `time` parameter, consult the help of `gtrends`. For instance, to get the results for the whole
history of Google trends, set `time = "all"`. For a specific time window, say 2020, set `time = "2020-01-01 2020-12-31"`.

Here is an example of an interesting "anti-correlation" for two keywords on a short time window in the US.
```{r}
#| label: example3
cor_tbl <- gtrends(keyword = c("ramen", "bitcoin"), geo = "US",time = "now 7-d")
cor_tbl$interest_over_time |> 
    mutate(hits = as.numeric(gsub("<", "", hits))) |> 
    ggplot(aes(x = date, y = hits, col = keyword))+
    geom_line()
```

